---
title: "EDLD_653_Lab3"
author: "Christopher Loan"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Part A: Foundations

## 1. Briefly name and describe the three fundamental components of a function.

* body — the code that's acting on (or under the influence of) the formals
* formals — the arguments supplied to the function; these control how the function is called
* environment — the location of the function's variables or the context in which the function acts. 

## 2. Describe three different ways functions can be stored/applied and an example of when you might want to use each version.

* Anonymous functions — something in `map()` and `*apply()` variants. Useful when you want to do something small that you won't do again
* User defined functions — make a function that you'll store in your local environment. Useful when you want to make a more complex function than anonymous functions, or if you need to integrate it into a larger/more complex process.
* Calling functions made by others — in a script or a package. Useful when you want to do something that's already been done.

# Part B: Applied practice
 
## 1. Write a function to calculate the mean that removes missing data before conducting the calculation. You may not use base::mean or any similar function that conducts a mean calculation. Include a warning in your function if missing data have been removed that prints the total number of cases that were removed.

```{r}
mean_function <- 
  function(test_vector){
    if (sum(is.na(test_vector) > 0)){
      missing_cases <- sum(is.na(test_vector))
      smaller_vector <- na.omit(test_vector)
      warning(
        'WARNING: Data contains ', 
        missing_cases, 
        ' missing cases; 
        value provided removes missing values.')
      return(sum(smaller_vector)/length(smaller_vector))
    } else{
      sum(test_vector)/length(test_vector)
      }
  }
```

## 2. Test your function to make sure it (a) provides the expected results, and (b) gives identical output to base::mean when na.rm = TRUE. Make sure that you test your data against a vector that has missing data.

```{r}
temp <- tibble(
  v1 = 1:20, 
  v2 = 21:40, 
  v3 = c(1:18, NA, NA), 
  v4 = letters[1:20]
  )

mean_function(temp$v1)
mean(temp$v1)
mean_function(temp$v2)
mean(temp$v2)
mean_function(temp$v3)
mean(temp$v3, na.rm = T)
```

## 3. Turn the following three lines of code into three different functions. Make sure to give them meaningful names. Test the functions to make sure they provide the expected output.

```{r}
missing_counter <- 
  function(test_vector){
    sum(is.na(test_vector))
  }
```

```{r}
proportional_rep <- 
  function(test_vector){
    purrr::map_dbl(
      split(test_vector, test_vector), length) / length(test_vector)
  }
```

```{r}
normalize_vector <- 
  function(test_vector){
    (test_vector - mean(test_vector, na.rm = TRUE)) / sd(test_vector, na.rm = TRUE) 
}
```

## 4. Write a function that takes a data frame as its input and returns a data frame with only the numeric columns.

```{r}
return_numeric_cols <- 
  function(df){
    index_vector <- 
      unlist(lapply(df, is.numeric))
    df[,index_vector]
  }
```

## 5. Write a function that uses the function you wrote in Question 5, and returns a data frame with the mean and standard deviation of each numeric column. If columns are omitted (because they are non-numeric), print a message stating the specific columns that were not included.

```{r}
means_sds <- 
  function(df){
    data.frame(
      rbind(
        mean = sapply(
          X = return_numeric_cols(df), 
          FUN = mean, 
          na.rm = T),
        sd = sapply(
          X = return_numeric_cols(df), 
          FUN = sd, 
          na.rm = T)
        )
      )
    }
```

```{r}
means_sds_robust <- 
  function(df){
    all_numeric <- 
      unlist(lapply(temp, is.numeric))
  
    if (sum(all_numeric == FALSE) > 0){
      message(sum(all_numeric == FALSE), 
              ' row(s) removed: ', 
              names(df[, !all_numeric])
              )
      return(means_sds(df))
  } else {
    means_sds(df)
    }
    }
means_sds_robust(temp)
```

