---
title: "EDLD_653_Lab3"
author: "Christopher Loan"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Part A: Foundations

## 1. Briefly name and describe the three fundamental components of a function.

## 2. Describe three different ways functions can be stored/applied and an example of when you might want to use each version.


# Part B: Applied practice
 
## 1. Write a function to calculate the mean that removes missing data before conducting the calculation. You may not use base::mean or any similar function that conducts a mean calculation. Include a warning in your function if missing data have been removed that prints the total number of cases that were removed.

```{r}
mean_function <- 
  function(test_vector){
    if (sum(is.na(test_vector) > 0)){
      missing_cases <- sum(is.na(test_vector))
      smaller_vector <- na.omit(test_vector)
      warning(
        'Data contains ', 
        missing_cases, 
        ' missing cases. Mean with NAs removed = ', 
        sum(smaller_vector)/length(smaller_vector)
      )
    } else{
      sum(test_vector)/length(test_vector)
      }
  }
```

## 2. Test your function to make sure it (a) provides the expected results, and (b) gives identical output to base::mean when na.rm = TRUE. Make sure that you test your data against a vector that has missing data.

```{r}
temp <- tibble(
  v1 = 1:20, 
  v2 = 21:40, 
  v3 = c(1:18, NA, NA)
  )

mean_function(temp$v1)
mean(temp$v1)
mean_function(temp$v2)
mean(temp$v2)
mean_function(temp$v3)
mean(temp$v3, na.rm = T)
```


## 3. Turn the following three lines of code into three different functions. Make sure to give them meaningful names. Test the functions to make sure they provide the expected output.

```{r}
missing_counter <- 
  function(test_vector){
    sum(is.na(test_vector))
  }
```

```{r}
proportional_rep <- 
  function(test_vector){
    purrr::map_dbl(
      split(test_vector, test_vector), length) / length(test_vector)
  }
```

```{r}
# normalize or z-score a vector so the mean is zero and sd is one
(mtcars$mpg - mean(mtcars$mpg, na.rm = TRUE)) / 
  sd(mtcars$mpg, na.rm = TRUE) 
```

## 4. Write a function that takes a data frame as its input and returns a data frame with only the numeric columns.

```{r}

```


## 5. Write a function that uses the function you wrote in Question 5, and returns a data frame with the mean and standard deviation of each numeric column. If columns are omitted (because they are non-numeric), print a message stating the specific columns that were not included.


